---
title: "iscience2023"
author: "Javad Nadaf"
date: "`r doc_date()`"
abstract: > 
 . Running seurat clustering on SVZ samples and selecting the NSC-cluster.<br>
 . Selecting progenitor cells from 2 fetal human brain datasets.<br>
 . Merging the selected cells from 3 datasets.<br>
 . Some visualizations
output:
  BiocStyle::html_document
---


## libraries

```{r}
source("functions_seurat.r")
library(pheatmap)
library(Seurat)
library(ggplot2)
library(ggplot2)
FINDALLMARKERS <- FALSE #  time consuming, run and save
```

## SVZ

### Run Seurat

```{r,eval=FALSE}
counts <- readRDS("counts_SVZ.rds")
meta.data <- readRDS("meta.data_SVZ.rds")
svzObj <- RunSeurat(counts = counts,meta.data = meta.data,nPCA = 35, resolution = 0.6)
```

```{r}
ElbowPlot(svzObj,ndims = 35)
JackStrawPlot(svzObj,reduction = "pca",dims = 1:35)
VlnPlot(svzObj,"percent.ribo")
VlnPlot(svzObj,"percent.mt")
```

### doublets predictions from scrublet

```{r}
pheatmap_w(table(svzObj$scrublet.predicted_doublets,svzObj$clusters_annotated),fontsize_number = 15)
```

### visualizing some markers

```{r}
DimPlot(svzObj,label = TRUE)
TSNEPlot(svzObj, label = TRUE)
svzObj@reductions$umap <- NULL
FeaturePlot(svzObj,features = c("PTPRZ1","BCAN"))
FeaturePlot(svzObj,features = c("SOX4","SOX11"))
FeaturePlot(svzObj,features = c("FABP7","EGFR"))
```

### Cell cycle scores

```{r}
FeaturePlot(svzObj,features = c("G2M.Score"),cols = cols.fp,pt.size=1)
FeaturePlot(svzObj,features = c("S.Score"),cols = cols.fp,pt.size=1)
```

### number of cells per samples and clusters

```{r}
x=as.matrix(table(svzObj$sample,svzObj$res.0.6))
pheatmap_w(x)


x=as.matrix(table(svzObj$sample,svzObj$clusters_annotated))
pheatmap_w(x)


```

### selecting NSC-like cells

```{r fig.height=5, fig.width=14}

idents <- "NSC-like"
sub1 <- subset(svzObj, idents = idents )
cells <- WhichCells(svzObj,idents = idents)
fig1 <- TSNEPlot(svzObj, label = TRUE)
fig2 <- TSNEPlot(svzObj,cells.highlight =  cells);fig2$data$highlight <-  plyr::mapvalues(fig2$data$highlight,"Group_1","Selected")
print(fig1+fig2)

if(FINDALLMARKERS){
  svzMk <- FindAllMarkers(svzObj)
}
```

## Fetal (Couturier 2020)

```{r,eval=FALSE}
counts <- readRDS("counts_Fetal.rds")
meta.data <- readRDS("meta.data_Fetal.rds")
meta.data$platform <- "seq10x"
fetalObj <- RunSeurat(counts = counts,meta.data = meta.data,nPCA = 30, resolution = 0.6)
```

```{r}
ElbowPlot(fetalObj,ndims = 30)
JackStrawPlot(fetalObj,reduction = "pca",dims = 1:30)
VlnPlot(fetalObj,"percent.ribo")
VlnPlot(fetalObj,"percent.mt")
```

```{r}
DimPlot(fetalObj, label = TRUE)
FeaturePlot(fetalObj,features = c("G2M.Score"),cols = cols.fp,pt.size=1)
FeaturePlot(fetalObj,features = c("S.Score"),cols = cols.fp,pt.size=1)
if(FINDALLMARKERS){
  fetalMk <- FindAllMarkers(fetalObj)
}
```

### selecting RG/progenitor clusters

```{r fig.height=5, fig.width=10}
idents <- c("RG","RG.Astro.Epend","RG.Astro","OPC")
sub2 <- subset(fetalObj, idents = idents )
cells <- WhichCells(fetalObj,idents = idents)
fig1 <- DimPlot(fetalObj, label = TRUE)
fig2 <- DimPlot(fetalObj,cells.highlight =  cells);fig2$data$highlight <-  plyr::mapvalues(fig2$data$highlight,"Group_1","Selected")
print(fig1+fig2)
```

## Fetal Nowakowski 2017, cells and clusters information

```{r fig.height=14, fig.width=14}
df=read.delim("sampleinfo_nowak.txt")
head(df)
x <- as.matrix(table(df$Name))
```

### selected clusters

```{r}
selected_clusters <- c("tRG","RG-div1","oRG","vRG","RG-div2","OPC","Astrocyte","RG-early")
```

## merging the 3 datasets

### selected SVZ and Fetal clusters

```{r,eval=FALSE}
counts <- readRDS("counts_sub123.rds")
meta.data <- readRDS("metadata_sub123.rds")
```

```{r}
table(meta.data$platform,meta.data$DataSet)
table(meta.data$orig.annotation)
```

### Run Seurat

```{r,eval=FALSE}
mergedObj <- RunSeurat(counts = counts,meta.data = meta.data,nPCA = 20, resolution = 0.8,vars.to.regress = "platform")
```

```{r}
ElbowPlot(mergedObj)
JackStrawPlot(mergedObj,reduction = "pca",dims = 1:20)
```

### visualization original annotations

```{r}
DimPlot(mergedObj,group.by = "orig.annotation",label = TRUE)
```

### visualization of Seurat clusters

```{r}
DimPlot(mergedObj, label = TRUE)
```

### visualizing some markers

```{r fig.height=5, fig.width=14}
FeaturePlot(mergedObj,features = c("EGFR","OLIG1"))
FeaturePlot(mergedObj,features = c("CKB","VIM"))
FeaturePlot(mergedObj,features = c("GFAP","SOX9"))
```

### add annotation

```{r}
clusters=c(4,5,7,9)
names(clusters)=c("npNSC","aRG1","aRG2","opNSC")
for(i in 1:length(clusters)){
cells = WhichCells(mergedObj,expression = seurat_clusters == clusters[i] & DataSet=="SVZ")
mergedObj@meta.data[cells,"annotation"] <- names(clusters)[i]
}
```

### visualization of SVZ cells annotations

```{r}
DimPlot(mergedObj,group.by = "annotation",label = TRUE)
if(FINDALLMARKERS){
  mergedMk <- FindAllMarkers(mergedObj)
}
```

### number of cells per annotated clusters

```{r}
x <- as.matrix(table(mergedObj$orig.annotation,mergedObj$seurat_clusters))
pheatmap_w(x)

x=as.matrix(table(mergedObj$sample,mergedObj$seurat_clusters))
pheatmap_w(x)

x=as.matrix(table(mergedObj$sample,mergedObj$annotation))
pheatmap_w(x)

```

### cell cycle scores and plots

```{r}
FeaturePlot(mergedObj,features = c("G2M.Score"),cols = cols.fp,pt.size=1)
FeaturePlot(mergedObj,features = c("S.Score"),cols = cols.fp,pt.size=1)
```

### Ribosomal protein genes are high in eRG and aRG1, with low transcriptome size and similar percentage of mitochondrial genes, within each dataset

```{r}
VlnPlot(subset(mergedObj,subset = DataSet == "Fetal_N"),"percent.ribo",group.by = "orig.annotation")
VlnPlot(subset(mergedObj,subset = annotation %in% c("aRG1","aRG2","opNSC","npNSC")),"percent.ribo",group.by = "annotation")

VlnPlot(subset(mergedObj,subset = DataSet == "Fetal_N"),"nFeature_RNA",group.by = "orig.annotation")
VlnPlot(subset_SVZ(mergedObj),"nFeature_RNA",group.by = "annotation")

VlnPlot(subset(mergedObj,subset = DataSet == "Fetal_N"),"percent.mt",group.by = "orig.annotation")
VlnPlot(subset(mergedObj,subset = annotation %in% c("aRG1","aRG2","opNSC","npNSC")),"percent.mt",group.by = "annotation")
```


## Markers expression visualization on dotplot (Figure 2C)

```{r fig.height=4, fig.width=14}

genes <- read.delim("data/genes_figure2C.txt")#fig 2C
DotPlot(svzObj,features = genes,)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size = 10))

ss <- subset_SVZ(mergedObj)
DotPlot(ss,features = genes,group.by = "annotation")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size = 10))

```
